Supplemental Information:\\ Harnessing Design Principles from Glass Sponges for Structurally Robust Lattice Structures

### 2018-10-03 Paper-FernandesWeaverBertoldi_AdvMaterials_2018.tex
.
katia's new edits

\documentclass[9pt,twocolumn,twoside]{fernandes_paper} 
%Default opts:9pt,twocolumn,twoside
\graphicspath{{images_paper/}}

@@

\begin{document}
\maketitle
% \doublespace
\linenumbers

% \section{Introduction}
% \subsection{Sponge Biology}
% \subsection{History of Diagonally Reinforced Structures and Previous Works}
% \subsection{Examples of Current Diagonally Reinforced Designs}

% \section{Methods}
% \subsection{Numerical Model Setup}
% \subsection{Experimental Design}

% \section{Results}
% \subsection{Linear Stiffness for different Designs}
%  \subsection{Buckling Resistance for Different Designs}
% \subsection{Experimental Results}

% \section{Discussion}
% \subsection{Parameter Exploration}
% \subsection{Parameter Optimization}

% \section{Conclusion}

% \section*{Acknowledgments}
% \section*{References}

@@

\begin{figure*}[ht]
	\centering
	\captionsetup{width=0.8\textwidth}
	\includegraphics[width=0.9\textwidth]{Fig2}
	\caption{text}\label{Fig2}
\end{figure*}

\begin{figure*}[ht]
	\captionsetup{width=0.8\textwidth}
	\begin{center}
		\includegraphics[width=0.7\textwidth]{Fig3}
	\end{center}
	\caption{\textbf{text} Figure showing (a) the structural stiffness for the different designs and (b) the buckling resistance for the different designs. The designs considered are depicted below as Design A-D.} \label{Fig3}
\end{figure*}

% \section{Introduction}

In modern day structural engineering, we find many

@@

Hexactinellids, also known as glass sponges, are predominately deep sea sponges that live in ocean depths of 100-2000m. Beyond their fracture propagation inhibiting material composition \citep{weaver2007}, they are perceived to exhibit large structural rigidity and strength against buckling. 
% Since they are primarily made of 'brittle silica', buckling strength may be a crucial property in making them resistant to impact and environmentally applied stresses. 
Structurally, they are composed of a base square-grid architecture and regular ordering of vertical and horizontal struts that form the skeletal system. Furthermore, their base structure is overlaid with double diagonal (two diagonals in each direction) reinforcement struts, which create a checkerboard-like pattern of open-closed cell structure. This diagonal reinforcement design is conjectured to give the sponge greater buckling resistance and strength to localized damage then it would experience having a single diagonal reinforcement strut (with equivalent allocation of mass between the diagonal reinforcements.) 
% Analogous to the sponge, many engineering structures, such as buildings and bridges, exhibit diagonal reinforcement struts as a stability mechanism. Based on this similarity, we explore the following research question:
The similarity in the structures of the sponge architecture and what we see in the Long truss design has inspired us to further explore the following research question: \textbf{Can we generate design principles for diagonal reinforcements of square beam lattices that are optimally designed to avoid global and/or local structural buckling?} \mf{this paragraph needs more citations and more avid lit. survey}

In this paper, we present a numerical analysis of the structure deformation under various loading conditions as well as survey different arrangements within similar design space of the sponge via parametric studies. Through the various design iterations we look for the critical buckling strain and the elastic load carrying capacity. Finally, we compare the results obtained from the sponge design to what is typically used in engineering structures and validate our results with experimental data.

% involving the development of truss structures with efficient diagonal reinforcement, is inspired by glass sponges (primarily made of silica) commonly known as the Venus Flower Basket. These deep sea sponges exhibit a regular and periodic truss structure which biologists for many years have stipulated to be an optimized design for structural and hydrodynamic purposes. To investigate these claims, I have built numerical and analytical models analyzing the buckling effects of the sponge designs by tuning various geometrical parameters and analyzing its effect on the buckling response. Thus far, we have found the the double diagonal reinforced sponge design delays the onset of structural buckling by a factor of twice the applied strain in comparison to the traditional designs. We believe that this intricate approach inspired by the sponge can improve the way we construct buildings and bridges, which will give the opportunity to build taller buildings and bridges that use less material and are stronger.

% \section{Results}
% \subsection{Linear Stiffness for different Designs}
In this section, we compare the linear stiffness for the different arrangements using Finite Element Analysis (FEA). For each design, we consider a Representative Volume Element (RVE) and apply periodic boundary conditions along the edges of the unit cell domains depicted along the bottom of \cref{Fig3}. In our numerical model we assume Timoshenko beam elements with uniform circular cross-sections and pin-joints at beam intersections. In an effort to create a method for impartial comparison, we maintain a constant total mass (or constant volume as the material is assumed to be incompressible) between all of the designs, as well as a constant mass ratio between diagonal and non-diagonal elements for designs containing these diagonal reinforcement.

@@

% \subsection{Buckling Resistance for Different Designs}
The results shown in \cref{Fig3}(b) indicate that the structural buckling strength is clearly dependent on the design and alignment of the beam elements. From this figure, we can see that Design A, the sponge design, has the highest buckling strength compared to the other designs containing diagonal reinforcement. At $45^\circ$ loading, Design A improves the buckling strength by a factor of 2.5 times that of Design C, which is the commonly used truss systems.

% \subsection{Experimental Validation}
For the experimental validation, we conducted various uniaxial compression tests of 3D printed planar beam structures (with circular cross-section) on an Instron 5566. The specimen was 3D printed using \hl{xxxx} material with manufacturer reported \hl{xxxx GPa} modulus. The printed specimen consisted of a 10x10 cell structure spanning a total of 10x10 inch planar structure. The experimental set-up consisted of two \hl{1 inch} thickness acrylic sheets bolted together along the perimeter and spaced from each other by the thickness of the 3D printed specimen.\\\\\hl{[...]} \mf{more to go here when we have experimental setup and data}

% \figll{Buckling_Separation}{0.9}{Critical buckling strain for varying diagonal separation in Design A parametric sweep. The normalized separation distance $C/L$ indicates how far apart the diagonals are from each other. For $C/L=0$ we obtain the Design B and for $C/L=0.5$ we obtain the diagonals crossing half way through the length of the non-diagonal trusses ($L$). For all values in this sweep, we assume a mass ratio equivalent to that of the sponge. The dashed gray line indicates the assumed diagonal separation of the sponge of $C/L=1/(\sqrt{2}+2)$.}

@@

% \section{Mass Ratio Parametrization}
% \figll{Stiffness_MassRatio}{0.9}{Structural stiffness for varying mass ratio parametric sweep. Mass ratio $\lambda$ indicates how much mass is allocated to the diagonal elements and non-diagonal elements. For all values in this graph, the design is kept constant while the cross-sectional area of each element is scaled accordingly. For all values of $\lambda$, the total mass of all elements are kept constant. The different colored lines correspond to different loading angles of the unit cell. The gray dashed line indicates the measured mass ratio for the sponge. For all values of this sweep, we assume Design A geometry although the linear stiffness does not depend on the design.}

% \figll{Buckling_MassRatio}{0.9}{Critical buckling strain for varying mass ratio parametric sweep. Mass ratio $\lambda$ indicates how much mass is allocated to the diagonal elements and non-diagonal elements. For all values in this graph, the design is kept constant while the cross-sectional area of each element is scaled accordingly. For all values of $\lambda$, the total mass of all elements are kept constant. The different colored lines correspond to different loading angles of the unit cell. The gray dashed line indicates the measured mass ratio for the sponge. For all values of this sweep, we assume Design A geometry.}

% \section{Diagonal Separation Parametrization}

@@

{Fig5}(b) shows the structural stiffness as a function of $\lambda$ for various loading angles (corresponding to the respective colors). In this figure, we see that for small $\lambda$ the structure is stiffer at $0^\circ$ loading and for large $\lambda$ the structure is stiffer at $45^\circ$. This is expected as we are allocating more material to its respective elements. Furthermore, we can see that at $\lambda=1$ we attain a point in the mass ratio where the stiffness is equal for all angles. That is because there is the same amount of material allocated to the diagonals as there is to the non-diagonal elements.

{Fig5}(c) illustrates the critical buckling strain as a function of $\lambda$. Here, we can clearly see that buckling strain is dependent on the mass ratio and behaves differently for all angles \hl{[...]} \mf{I am not sure what to say here (interpretation) about this graph.}

% \section{Conclusion}
% \hl{add writing}

\section*{Funding Information}
\hl{add writing}
% No. NSF-GRFP DGE-1144152 (for M. C. F.), GEM Consortium Fellowship

\section*{Acknowledgments}
\hl{add writing} 
% Wyss Institute for equipment. James R. Rice. Mark Scott. Joanna Aizenberg. Joost Vlassak. Chris Rycroft

%\begin{figure*}
%	\captionsetup{width=0.8\textwidth}
%	\begin{center}
%		\includegraphics[width=0.7\textwidth]{Fig5}
%	\end{center}
%	\caption{\textbf{Parametric sweep results for different analysis.} (a) Critical buckling strain for varying diagonal separation in Design A. The normalized separation distance $C/L$ indicates how far apart the diagonals are from each other. For all values we assume a mass ratio to be equivalent to that of the sponge. The dashed gray line indicates the assumed diagonal separation of the sponge, namely $C/L=1/(\sqrt{2}+2)$. (b) Structural linear stiffness for varying mass ratio. Mass ratio $\lambda$ indicates how much mass is allocated to the diagonal elements and non-diagonal elements based on the volume element. For all values of $\lambda$, the total mass of all elements are kept constant. The gray dashed line indicates the measured mass ratio for the sponge. (c) Critical buckling strain for varying mass ratio. For all values of $\lambda$, the total mass of all elements are kept constant. The different colored lines correspond to different loading angles of the unit cell. The gray dashed line indicates the measured mass ratio for the sponge. For all values in (b) and (c), we assume Design A geometry.} \label{Fig5}
%\end{figure*}

 Bibliography
\nocite{*}
\bibliography{refs}
\bibliographystyle{apalike}

>>>>

\documentclass[10pt,twocolumn,twoside]{fernandes_paper} 
%Default opts:9pt,twocolumn,twoside
\graphicspath{{images_paper/}}

@@

\begin{document}
\maketitle
% \doublespace
%\linenumbers

@@

\begin{figure*}[ht]
	\centering
	\captionsetup{width=0.8\textwidth}
	\includegraphics[width=0.9\textwidth]{Fig2}
	\caption{\textbf{Experimental Results.} (a) shows the different designs (Design A-D) considered in this analysis. (b) shows the undeformed experimental setup for the different designs. (c) shoes the post-buckling deformed state for each of the designs. (d) shows the experimental stress-strain curve as well as the overlaid numerical results for linear stiffness and critical buckling for the different designs considered.}\label{Fig2}
\end{figure*}

\begin{figure*}[ht]
	\captionsetup{width=0.8\textwidth}
	\begin{center}
		\includegraphics[width=0.7\textwidth]{Fig3}
	\end{center}
	\caption{\textbf{Structure Mechanical Response} Figure showing (a) the structural stiffness for the different designs and (b) the buckling resistance for the different designs. The designs considered are depicted below as Design A-D.} \label{Fig3}
\end{figure*}

\begin{figure*}[ht]
	\captionsetup{width=0.8\textwidth}
	\begin{center}
		\includegraphics[width=0.7\textwidth]{Fig4}
	\end{center}
	\caption{\textbf{Critical Buckling Load Optimization Results.} (a) shows the optimal value of critical buckling load for varying number of diagonals. The color of each point represents the optimal mass ratio $\lambda$  parameter for that configuration.  Note, we do not explicitly show the other optimal values for the diagonal spacing parameters. (b) shows the resulting deformed geometries for designs including one to four diagonals.} \label{Fig4}
\end{figure*}

@@

\textit{Hexactinellids}, also known as glass sponges, are predominately deep sea sponges that live in ocean depths of 100-2000m. Beyond their fracture propagation inhibiting material composition \citep{weaver2007}, they are perceived to exhibit large structural rigidity and strength against buckling.

Structurally, they are composed of a base square-grid architecture and regular ordering of vertical and horizontal struts that form the skeletal system. Furthermore, their base structure is overlaid with double diagonal (two diagonals in each direction) reinforcement struts, which create a checkerboard-like pattern of open-closed cell structure. This diagonal reinforcement design is conjectured to give the sponge greater buckling resistance and strength to localized damage then it would experience having a single diagonal reinforcement strut (with equivalent allocation of mass between the diagonal reinforcements.)

The similarity in the structures of the sponge architecture and what we see in the Long truss design has inspired us to further explore the following research question: \textbf{Can we generate design principles for diagonal reinforcements of square beam lattices that are optimally designed to avoid global and/or local structural buckling?}

In this paper, we present a numerical and experimental mechanical analysis of multiple  structures under various loading conditions as well as survey different arrangements within similar design space via parametric studies. Furthermore, we incorporate an optimization algorithm on our parameter space to obtain an optimal design to withstand (via critical buckling load) uniaxial loading parallel to the non-diagonal truss members. For this analysis, we restrict ourselves to four designs considered (Design A-D) that are shown along the top of \cref{Fig2}. As part of our analysis, we study the critical buckling strain and the elastic load carrying capacity of these structures.

Numerically, we compare the linear stiffness and critical buckling strain for each of the different designs using Finite Element Analysis (FEA). To reduce the computational cost, we took advantage of the periodicity of the structures and investigated their response  by considering a Representative Volume Element (RVE) with periodic boundary conditions along the free edges of the unit cell (more information on this can be found in the SI). For our numerical model we assume Timoshenko beam elements with uniform circular cross-sections and pin-joints at beam intersections. In an effort to create a method for impartial comparison, we maintain a constant total mass (or constant volume as the material is assumed to be incompressible) between all of the designs, as well as a constant mass ratio between diagonal and non-diagonal elements for designs containing these diagonal reinforcement.

@@

The results shown in \cref{Fig3}(b) indicate that the structural buckling strength is clearly dependent on the design and alignment of the beam elements. From this figure, we can see that Design A, the sponge design, has the highest buckling strength compared to the other designs containing diagonal reinforcement. At $45^\circ$ loading, Design A improves the buckling strength by a factor of 2.5 times that of Design C, which is the commonly used truss systems.

For the experimental validation, we conducted various uniaxial compression tests of 3D printed planar beam structures (with circular cross-section) on an Instron 5566. The specimen was 3D printed using \hl{xxxx} material with manufacturer reported \hl{xxxx GPa} modulus. The printed specimen consisted of a 10x10 cell structure spanning a total of 10x10 inch planar structure. The experimental set-up consisted of two \hl{1 inch} thickness acrylic sheets bolted together along the perimeter and spaced from each other by the thickness of the 3D printed specimen.\\\\\hl{[...]} \mf{more to go here when we have experimental setup and data}

@@

@@

%\section*{Acknowledgments}

% No. NSF-GRFP DGE-1144152 (for M. C. F.), GEM Consortium Fellowship
% Wyss Institute for equipment. James R. Rice. Mark Scott. Joanna Aizenberg. Joost Vlassak. Chris Rycroft

\nocite{*}
\bibliography{refs}
\bibliographystyle{apalike}

### 2018-10-03 Supplemental-FernandesWeaverBertoldi_AdvMaterials_2018.tex
.

\section{Our four designs}\label{sec:designs}
In this study, we focus on four different lattice configurations (Designs A, B, C and D) constrained to deform in an in-plane setting only. In an effort to conduct a fair performance comparison between the different designs, all four designs are characterized by the same total volume of material and a fixed volume ratio between non-diagonal and diagonal elements (chosen to match the sponge geometry).  Two different shapes are considered for teh cross section of the struts: circular and rectangular.  For the circular cross-section case  we denote with $D_{\alpha,nd}$ and $D_{\alpha,d}$ the diameter of the non-diagonal (i.e. horizontal and vertical) and diagonal struts in the $\alpha$-th design, respectively, and neglect out-of-plane buckling. For the rectangular cross-sections we choose the depth $H$ and in-plane thickness $T_{\alpha,nd}$ and $T_{\alpha,d}$  to avoid out-of-plane deformation (i.e. we choose the  depth over thickness ratio sufficiently large). Finally, it is important to note that teh slenderness of the non-diagonal members in the $\alpha$-th design is chosen as 
\begin{equation}
\frac{D_{\alpha,nd}}{L}=0.1,\;\;\;\text{and}\;\;\;\frac{T_{\alpha,nd}}{L}=0.1,
\end{equation}
for the case of circular and rectangular cross section, since this is the aspect ratio measured for the sponges (see \cref{sec:params}).

In the subsequent sections, we describe in detail the unit cells for four different designs (Designs A-D), and provide the derivations for each geometry cross-section characteristics.

\subsection{Design A}
Design A is inspired by the sponge structure and consists of a square grid  reinforced by a double diagonal system (see \cref{DesignA}). As for the case of the sponge, the diagonal elements are assumed to form an octagonal opening on every other cell, so that they  intersect the horizontal and vertical struts at a distance $\Delta L=L/(\sqrt{2}+2)$ from the nodes, where $L$ denote the length of the vertical/horizontal struts. 
\subsubsection{Circular cross section} 
If we assume that cross section of all struts is circular, the projected-area and volume for the non-diagonal ($A_{A,nd}$ and $V_{A,nd}$) and diagonal ($A_{A,d}$ and $V_{A,d}$) members is given by
\begin{equation}\label{A1}
A_{A,nd}=8 L D_{A,nd},
\end{equation}
\begin{equation}\label{V1}
V_{A,nd}=8 L \left(\pi\frac{D^2_{A,nd}}{4}\right)=2 L \pi D^2_{A,nd}
\end{equation}
\begin{equation}\label{A2}
A_{A,d}=8\sqrt{2} L D_{A,d},
\end{equation}
and
\begin{equation}\label{V2}
V_{A,d}=8 \sqrt{2} L\left(\pi\frac{D^2_{A,d}}{4}\right)=2\sqrt{2}L \pi D^2_{A,d}.
\end{equation}

@@

Finally, also for the case of rectangular cross-section we use Design A as our base model, and thus constrain the total volume of all the other unit cell designs with rectangular cross-section to be equal to that of Design A, namely, \KB{update eqn}
\begin{equation}\label{con1R}
V_{\alpha,d}+V_{\alpha,nd}={V}_{A,T}= ??,
\end{equation}
with $\alpha=$ B, C and D.
Moreover, for Designs B and C, which comprise diagonal elements, we also  constrain the volume ratio of the non-diagonal to diagonal elements to be the same as in Design A \KB{update eqn}
\begin{equation}\label{con2R}
\frac{V_{\alpha,nd}}{V_{\alpha,d}}=???,
\end{equation}
with $\alpha=$ B and C.

@@

\subsubsection{Rectangular cross section}
For this design with circular cross section the volume of the non-diagonal and diagonal mmebers  are given by
\begin{equation}
		V_{B,nd}=8LT_{B,nd}H.
\end{equation}
and
\begin{equation}
		V_{B,d}=4\sqrt{2}LT_{A,d}H.
\end{equation}
Using the constraints provided by \cref{con1R} and \cref{con2R}, as well as the above volumes, we  obtain
\begin{equation}
	T_{B,nd}=T_{B,d}
\end{equation}
and
\begin{equation}
	T_{B,nd}=T_{A,nd}
\end{equation}

@@

\subsubsection{Rectangular cross section}
For this design with circular cross section, the volume of the non-diagonal and diagonal members  of the unit cell are given by
\KB{add eqns.}
Using the constraints provided by \cref{con1R} and \cref{con2R}, as well as the above volumes, we  obtain
\begin{equation}
T_{A,nd}=2T_{A,d},
\end{equation}
and
\begin{equation}
T_{C,nd}=T_{A,nd}.
\end{equation}

@@

To reduce the computational cost, we took advantage of the periodicity of the structures and investigated their response  using  the unit cells shown  in \cref{RVE}.
To subject the unit cells to a macroscopic deformation gradient $\overline{\mathbf{F}}$ periodic boundary conditions are imposed on all cell boundaries by enforcing  \citep{danielsson2002, bertoldi2008}
\begin{equation} \label{EQN:DefInfPeriodicFEM}
    u_\alpha^{A_i} -u_\alpha^{B_i} = (\overline{F}_{\alpha\beta}-\delta_{\alpha\beta})(X_\beta^{A_i}-X_\beta^{B_i}), \quad i=1,\,2,\,....,\,N
\end{equation}
where $\delta_{\alpha\beta}$ is the Kronecker delta, $u_\alpha^{A_i}$  and $u_\alpha^{B_i}$ ($\alpha=1,\,2$) are displacements of points periodically located on the boundary of the unit cell. Moreover, $N$ denotes the number of pairs of nodes periodically located on the boundary of the unit cell. \KB{what does the following sentence mean? Are you referring to the unit cells shown in Fig. 6?}For the square lattice considered in this study $N=2$ when the unit cell is considered, while $N=4$ and $N=6$ for the enlarged $2\times2$ and $1\times5$ unit cells, respectively. Note that  the components of $\overline{\mathbf{F}}$ can be conveniently prescribed within the finite element framework using a set of virtual nodes. The corresponding  macroscopic first Piola-Kirchoff stress $\overline{\mathbf{P}}$ is then obtained through virtual work considerations \citep{danielsson2002, bertoldi2008}.

To subject the structures to uniaxial compression, in this study the following uniaxial macroscopic loading condition was considered:\KB{are you sure? I believe we left $\lambda_x$ unset. Right?} \begin{equation}
\overline{\mathbf{F}} = \begin{bmatrix}
0&0\\
0&\lambda_y
\end{bmatrix} 
\end{equation}

@@

\begin{figure}
	\centering
	\includegraphics[width=0.25\linewidth]{SFig11.png}
	\caption{{\bf Schematic depicting parameters of Optimization.} This schematic shows an example optimization geometry  at a complexity level of five diagonals. For each set of even diagonals, we allow each diagonal to travel away from each other from the center junction with distances $S_1,S_2$. Note that for every odd diagonal, the distance $S_0=0$ and it is not allowed to vary in order to ensure symmetry.}
	\label{OptimizationSchematic}
\end{figure}

@@

The algorithm's initial values are chosen to be in the center of the design space, namely, $\lambda=1$ and \KB{is thsi $S_i$??}diagonal separation$=0.5*L$. The covariance matrix is initialized uniformly with standard deviation ($\sigma$) half of the domain space, which are normalized to remain between 0 and 1. The optimization is run for uniaxial loading condition in the direction parallel to the vertical elements.

@@

\subsection{Rectangular Cross Section}\label{sec:rectparam}
This section shows the results when using a rectangular cross-section for the truss members. From \cref{SquareCrossSectionParameter}(a) it is apparent that there exists an optimum for the diagonal separation that occurs when the spacing between diagonals are approx 0.2 of the horizontal distance between vertical struts. This optimum value also persists when varying the loading angle. From \cref{SquareCrossSectionParameter}(b) it can be seen that the linear stiffness is symmetrically and almost purely dependent on the mass ratio allocated to diagonal versus non-diagonal elements. Comparing this figure to \cref{CircularCrossSectionParameter}(b) we can also see that even the design cross-section does not change the linear stiffness behavior.  \cref{SquareCrossSectionParameter}(c) shows that there exists two optimum mass ratios, one where more material is allocated to the diagonal and one where there is more material allocated to non-diagonals.  Note that if more matterialk

@@

\section{Local and Global Instabilities}
\mf{How should I go about describing this here?}

% Bibliography
%\nocite{aizenberg2005}
%\nocite{deshpande2001}
%\nocite{miserez2008}
%\nocite{weaver2010}

>>>>

\section{Our four designs}\label{sec:designs}
In this study, we focus on four different lattice configurations (Designs A, B, C and D) constrained to deform in an in-plane setting only. In an effort to conduct a fair performance comparison between the different designs, all four designs are characterized by the same total volume of material and a fixed volume ratio between non-diagonal and diagonal elements (chosen to match the sponge geometry).  Two different shapes are considered for the cross-section of the struts: circular and rectangular.  For the circular cross-section case  we denote the diameters with $D_{\alpha,nd}$ and $D_{\alpha,d}$ of the non-diagonal (i.e. horizontal and vertical) and diagonal struts in the $\alpha$-th design, respectively, and neglect out-of-plane buckling. For the rectangular cross-sections we choose the depth $H$ and in-plane thickness $T_{\alpha,nd}$ and $T_{\alpha,d}$  to avoid out-of-plane deformation (i.e. we choose the  depth over thickness ratio sufficiently large). Finally, it is important to note that the slenderness of the non-diagonal members in the $\alpha$-th design $\in [A,B,C]$ is chosen as 
\begin{equation}
\frac{D_{\alpha,nd}}{L}=0.1,\quad\text{and}\quad\frac{T_{\alpha,nd}}{L}=0.1,
\end{equation}
for the case of circular and rectangular cross section, since this is the aspect ratio measured for the sponges (see \cref{sec:params}).

In the subsequent sections, we describe in detail the unit cells for four different designs, and provide the derivations for each geometry cross-section characteristics.

\subsection{Design A}
Design A is inspired by the sponge structure and consists of a square grid  reinforced by a double diagonal system (see \cref{DesignA}). As for the case of the sponge, the diagonal elements are assumed to form an octagonal opening on every other cell, so that they  intersect the horizontal and vertical struts at a distance $\Delta L=L/(\sqrt{2}+2)$ from the nodes, where $L$ denote the length of the vertical/horizontal struts. 
\subsubsection{Circular cross section} 
If we assume that cross section of all struts is circular, the projected area and volume for the non-diagonal ($A_{A,nd}$ and $V_{A,nd}$) and diagonal ($A_{A,d}$ and $V_{A,d}$) members is given by
\begin{equation}\label{A1}
A_{A,nd}=8 L D_{A,nd},
\end{equation}
\begin{equation}\label{V1}
V_{A,nd}=8 L \left(\pi\frac{D^2_{A,nd}}{4}\right)=2 L \pi D^2_{A,nd}
\end{equation}
\begin{equation}\label{A2}
A_{A,d}=8\sqrt{2} L D_{A,d},
\end{equation}
and
\begin{equation}\label{V2}
V_{A,d}=8 \sqrt{2} L\left(\pi\frac{D^2_{A,d}}{4}\right)=2\sqrt{2}L \pi D^2_{A,d}.
\end{equation}

@@

Finally, also for the case of rectangular cross-section we use Design A as our base model, and thus constrain the total volume of all the other unit cell designs with rectangular cross-section to be equal to that of Design A, namely,
\begin{equation}\label{con1R}
{V}_{A,T}=V_{\alpha,d}+V_{\alpha,nd}= 8LH(T_{A,nd}+\sqrt{2}T_{A,d})=8LHT_{A,nd}\left(1+\frac{1}{\sqrt{2}}\right),
\end{equation}
with $\alpha=$ B, C and D.
Moreover, for Designs B and C, which comprise diagonal elements, we also  constrain the volume ratio of the non-diagonal to diagonal elements to be the same as in Design A,
\begin{equation}\label{con2R}
\frac{V_{\alpha,nd}}{V_{\alpha,d}}=\sqrt{2},
\end{equation}
with $\alpha\in$ B and C.

@@

\subsubsection{Rectangular cross section}
For this design with circular cross section the volume of the non-diagonal and diagonal members  are given by
\begin{equation}
		V_{B,nd}=8LT_{B,nd}H.
\end{equation}
and
\begin{equation}
		V_{B,d}=4\sqrt{2}LT_{B,d}H.
\end{equation}
Using the constraints provided by \cref{con1R} and \cref{con2R}, as well as the above volumes, we  obtain
\begin{equation}
	T_{B,nd}=T_{B,d}
\end{equation}
and
\begin{equation}
	T_{B,nd}=T_{A,nd}
\end{equation}

@@

\subsubsection{Rectangular cross section}
For this design with circular cross section, the volume of the non-diagonal and diagonal members  of the unit cell are given by
\begin{equation}
V_{C,nd}=8LT_{C,nd}H
\end{equation}
and
\begin{equation}
V_{C,d}=8\sqrt{2}LT_{C,d}H
\end{equation}
Using the constraints provided by \cref{con1R} and \cref{con2R}, as well as the above volumes, we  obtain
\begin{equation}
T_{C,nd}=2T_{C,d},
\end{equation}
and
\begin{equation}
T_{C,nd}=T_{A,nd}.
\end{equation}

@@

To reduce the computational cost, we took advantage of the periodicity of the structures and investigated their response  using  the unit cells shown  in \cref{RVE}.
To subject the unit cells to a macroscopic deformation gradient $\overline{\mathbf{F}}$ periodic boundary conditions are imposed on all cell boundaries by enforcing  \citep{danielsson2002, bertoldi2008}
\begin{equation} \label{EQN:DefInfPeriodicFEM}
    \mathbf{u}_\alpha^{A_i} -\mathbf{u}_\alpha^{B_i} = (\mathbf{\overline{F}}_{\alpha\beta}-\delta_{\alpha\beta})(\mathbf{X}_\beta^{A_i}-\mathbf{X}_\beta^{B_i}), \quad i=1,\,2,\,....,\,K
\end{equation}
where $\delta_{\alpha\beta}$ is the Kronecker delta, $\mathbf{u}_\alpha^{A_i}$  and $\mathbf{u}_\alpha^{B_i}$ ($\alpha=1,\,2$) are displacements of points periodically located on the boundary of the unit cell. Moreover, $K$ denotes the number of pairs of nodes periodically located on the boundary of the unit cell.  Note that  the components of $\overline{\mathbf{F}}$ can be conveniently prescribed within the finite element framework using a set of virtual nodes. The corresponding  macroscopic first Piola-Kirchoff stress $\overline{\mathbf{P}}$ is then obtained through virtual work considerations \citep{danielsson2002, bertoldi2008}.

To subject the structures to uniaxial compression, in this study the following uniaxial macroscopic loading condition was considered: \begin{equation}
\overline{\mathbf{F}} = \begin{bmatrix}
\text{UNSET}&0\\
0&\lambda_y
\end{bmatrix} 
\end{equation}

@@

\begin{figure}
	\centering
	\includegraphics[width=0.2\linewidth]{SFig11.png}
	\caption{{\bf Schematic depicting parameters of Optimization.} This schematic shows an example optimization geometry  at a complexity level of five diagonals. For each set of even diagonals, we allow each diagonal to travel away from each other from the center junction with distances $S_1,S_2$. Note that for every odd diagonal, the distance $S_0=0$ and it is not allowed to vary in order to ensure symmetry.}
	\label{OptimizationSchematic}
\end{figure}

@@

The algorithm's initial values are chosen to be in the center of the design space, namely, $\lambda=1$ and diagonal separation for the even set of diagonals $S_i=0.5*L$. The covariance matrix is initialized uniformly with standard deviation ($\sigma$) half of the domain space, which are normalized to remain between 0 and 1. The optimization is run for uniaxial loading condition in the direction parallel to the vertical elements.

@@

\subsection{Rectangular Cross Section}\label{sec:rectparam}
This section shows the results when using a rectangular cross-section for the truss members. From \cref{SquareCrossSectionParameter}(a) it is apparent that there exists an optimum for the diagonal separation that occurs when the spacing between diagonals are approx 0.2 of the horizontal distance between vertical struts. This optimum value also persists when varying the loading angle. From \cref{SquareCrossSectionParameter}(b) it can be seen that the linear stiffness is symmetrically and almost purely dependent on the mass ratio allocated to diagonal versus non-diagonal elements. Comparing this figure to \cref{CircularCrossSectionParameter}(b) we can also see that even the design cross-section does not change the linear stiffness behavior.  \cref{SquareCrossSectionParameter}(c) shows that there exists two optimum mass ratios, one where more material is allocated to the diagonal and one where there is more material allocated to non-diagonals.

@@

\section{Local and Global Instabilities}
\mf{How should I go about describing this here?}
\begin{figure}
	\centering
	\includegraphics[width=0.7\linewidth]{SFig12.pdf}
	\caption{{\bf Schematic depicting parameters of Optimization.} This schematic shows an example optimization geometry  at a complexity level of five diagonals. For each set of even diagonals, we allow each diagonal to travel away from each other from the center junction with distances $S_1,S_2$. Note that for every odd diagonal, the distance $S_0=0$ and it is not allowed to vary in order to ensure symmetry.}
	\label{DomainSizeSchematic}
\end{figure}
% Bibliography
%\nocite{aizenberg2005}
%\nocite{deshpande2001}
%\nocite{miserez2008}
%\nocite{weaver2010}
